### 2015-07-07 narrative/placement.tex
restructured the entire life cycle text. Purging is still incomplete. Someone need to sanity check the migration portions also. We also need milestones for each of the subareas here

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
The overall data life cycle consists of three phases. Data starts
when it is generated by a simulation or through a data acquisition system,
in the case of experimental and observational data (EOD). 
In our target heterogenous multi-tier storage platform, the system must
decide where to place the output data at this time, and how to partition and
reorganize the data to optimize this initial placement. The key optimization
metrics during this phase is the percieved I/O overhead for the application.

@@

% In the final phase the data is purged from the system altogether. The
% decision to purge is made by the system based on operational metrics from
% system administrators. This phase optimizes the space available in the
% system to prevent out of space concerns for applications.

@@

 % Next we provide details on the challenges and proposed research through each
 % phase of the data life cycle.

@@

\paragraph{Data Utility:} We will formulate the concept of utility (from an
application perspective) that is assigned data objects to quantify the
benefits of placement and/or movement actions to application. These
utilities may be based on the importance of the data objects, the frequency
with which they are accessed, etc., and are used during runtime data
management. For example, data objects will higher utility value may be
placed closer (e.g., in faster memory) to the accessing application.
Similarly, when it is time to evict data objects from a higher level storage
layer, the decision about which object to evict may take the utility into
account.

\paragraph{Autonomic in-transit data management:} 
Efficiently placing data objects
vertically across multiple memory hierarchy layers and storage levels as
well as horizontally across different storage nodes requires balancing
competing objectives. We will explore autonomic data
management strategies that can evaluate utility/cost tradeoffs and
appropriately place/move data objects at runtime. The autonomic approach
will leverage both user hints and information gathered at runtime (e.g.,
runtime data access history, network topology, etc.) to place data object.

@@

% \paragraph{State of the art:} NAND flash-based devices are being
% increasingly used in HPC storage systems at different levels and for a
% variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
% have investigated using compute node local SSDs as storage buffers, for
% example, to temporarily cache checkpoint data to support recovery from
% failures. Research efforts have also explored using deep memory devices for
% data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
% data analysis by directly executing data analysis tasks on emerging storage
% devices. Minerva \cite{minerva} extends the conventional SSD architecture
% using a FPGA-based storage controller to offload data or I/O intensive
% application code to the SSD to accelerate data analysis. However, these
% solutions cannot deal with coupled application workflows, where the data
% exchange and access patterns can be both complex and dynamic. Furthermore,
% some of these solutions require hardware modifications or special access
% priorities to the HPC systems, which can limit their use to specific
% resources.

\subsubsection{Managing Data on Multiple Tiers}
\label{sec:manag-data-mult}

\paragraph{Background:} There is a spectrum of approaches on how
to leverage a memory and storage hierarchy: On one end of the
spectrum is to copy or move data from one tier to another, the
closer the tier to the hierarchy's top, the faster the access to
the data it contains. On the other end of the spectrum is to use
higher tiers exclusively for auxiliary data that to some extent
represents the actual data on the lowest tier. Examples of auxiliary
data are views, indices, lossy compressions, lower resolution data,
and summary data (sometimes also known as metadata). There are a
number of drivers that push approaches to one or the other extreme:
(1) if a high tier has sufficient room for an application's working
set, approaches that move the working set into that tier will perform
well; Conversely, if a high tier very rarely captures working sets,
it is better to fill it with auxiliary data that speeds up access
to the working set captured by a lower tier; working sets only
reduce capacity misses but the concept can be generalized to the
likelihood of having the right data in place, i.e. proactively
moving data to high tiers accessed in predictable patterns reduces
compulsory misses as well; (2) if the application generates parallel
read/write requests and strong consistency semantics are important,
approaches that use auxiliary data to speed up access to persistent
shared data might yield better performance than approaches that
require coherence overhead to keep all copies on higher tiers
consistent; (3) if the application's access patterns are mismatched
with low tier access characteristics, a combination of approaches
that use auxiliary data and move data between tiers in order to
convert access patterns into a better match will perform well.

% Any other drivers?
\paragraph{Thinning of Data:} 
This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so.

@@

\subsubsection{Migration}
\label{sec:migration}

\paragraph{Background:} One important research question in this proposal is
whether and how a liberated Hierarchical Storage Manager designed for online
use, such as Sirocco, might be employed in supporting the higher level
concepts discussed previously. Conceptually, in this effort, data is
accessed by the middleware I/O libraries at various tiers in the lower-level
storage hierarchy. While Sirocco does autonomously group the distributed set
of media into like groups and manage movement of data between these groups
the various motivations to do so include both the traditional ones found in
classical HSM (Hierarchical Storage Management) as well as the novel.

Sirocco groups like media into aggregated volumes on a service node, and like
service nodes with like media into pseudo-tiers. It accomplishes both using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

Grouping into pseudo-tiers is not a formal, nor rigid, concept within Sirocco,
at present. In classic HSM the tiering is often a formally expressed concept,
embodied both in the architecture and the implementation. In Sirocco, though,
it is not directly expressed. Instead, a client or server that attempts to make
a decision relative to migration and staging uses the attributes discussed
above to constrain the choice of candidates. For instance, a server acting as
an off-node writeback cache for a compute node might only have access to
volatile dynamic RAM for use as storage media and a client might deposit data
that is marked for persistent media as a resilience constraint. At some point,
then, such a server must copy or move the data to another Sirocco server with
media matching, or exceeding, the resilience constraint. At that point in time,
all servers managing appropriate persistent media are candidates but
second-order attributes, such as latency and bandwidth of a candidate's managed
media would tend to dominate the ranking. Thus, the concept of a ``tier''
becomes an artifact of decision rather than a formal concept.

Further complication arises when a Sirocco server is forced to make choices for
capacity reasons. The Sirocco architecture dictates that a server function as a
victim cache. As media managed by a server fills the server will begin to
manage for capacity. In order to do so, currently, it need only verify or make
a copy of the data on another server that meets the resiliency constraint of
the held data prior to removing the local copy and in the current
implementation the server makes no attempt either to remember it ever had the
data or to notify others about what it is doing.  others that it has removed
the local copy or remember that it has

Taken together then, the above at least complicates the higher level research
questions discussed in this proposal, potentially even renders some
unachievable since a few involve determinism. We are forced, then, to consider
either or both mitigating management policies within and between Sirocco
servers as well as explicitly allowing clients to participate in capacity
management decisions, for a time at least, on the relevant Sirocco servers.

\paragraph{Bounded estimates for access:}
Our research intent involves short term and long term activities. In the short
term we will modify the relevant clients to provide hit and miss related data
and associated activities and explore an enhancement to our storage servers
that would provide some, hopefully tightly, bounded estimate about time to
first byte for a potential read operation.

\paragraph{Migration Policies:}
In the long term we hope that an
augmented client-server API providing usage hints will allow us to modify local
server policy algorithms to more effectively manage it's media and communicate
to other servers holding copies of the relevant data how they might best manage
theirs. In the event that this approach proves ineffective at meeting the
overall approach described in this proposal we will explore modifying Sirocco
servers to provide notifications about locations of the relevant data as they
migrate and stage as well as supporting client-directed pinning of location,
for a period at least. Client-directed pinning could introduce potentially
serious negative impacts to Sirocco's overarching goals though, since the
fundamental design of Sirocco assumes much about opaque and unrestricted data
movement, so this is contemplated only as a last resort.

\subsubsection{Preliminary work and results} 
\label{sec:prework}

Our recent work \cite{tongipdps15}
explored a two-tiered staging method that spans both DRAM and solid state
disks (SSD). It allows us to support both code coupling and data management
for data intensive simulation workflows in a loosely coupled manner. In
addition, our application-aware adaptive data placement has demonstrated the
effectiveness of using user provided access information as ``hints'', and
runtime data access pattern history to improve data access efficiency and
overall end-to-end execution. Besides, additional data placement
optimization \cite{qiansc15} that leverages physical network topology along
with data access patterns has also demonstrated lower data access costs and
good scalability. Our work have been deployed and proven with real
applications, e.g., coupled combustion simulations (DNS-LES) that are part
of the ExaCT Exascale Co-design Center and coupled Fusion simulations
(XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems
such Titan at ORNL.

>>>>

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
The overall data life cycle consists of four phases. Data starts
when it is generated by a simulation or through a data acquisition system,
in the case of experimental and observational data (EOD). 
In our target heterogenous multi-tier storage platform, the system must
decide where to place the output data at this time, and how to partition and
reorganize the data to optimize this initial placement. The key optimization
metrics during this phase is the percieved I/O overhead for the application.

@@

In the final phase the data is purged from the system altogether. The
decision to purge is made by the system based on operational metrics from
system administrators. This phase optimizes the space available in the
system to prevent out of space concerns for applications.

@@

%Next we provide details on the challenges and proposed research through each
% phase of the data life cycle.

@@

\paragraph{Data Utility:} We will formulate the concept of utility (from an
application perspective) that is assigned data objects to quantify the
benefits of placement and/or movement actions to application. These
utilities may be based on the importance of the data objects, the frequency
with which they are accessed, etc., and are used during runtime data
management. For example, data objects will higher utility value may be
placed closer (e.g., in faster memory) to the accessing application.
Similarly, when it is time to evict data objects from a higher level storage
layer, the decision about which object to evict may take the utility into
account. We will also explore the balancing of the data across the memory
hierarchy autonomically. The autonomic approach will leverage both user
hints and information gathered at runtime (e.g., runtime data access
history, network topology, etc.) to place data object.

@@

% \paragraph{State of the art:} NAND flash-based devices are being
% increasingly used in HPC storage systems at different levels and for a
% variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
% have investigated using compute node local SSDs as storage buffers, for
% example, to temporarily cache checkpoint data to support recovery from
% failures. Research efforts have also explored using deep memory devices for
% data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
% data analysis by directly executing data analysis tasks on emerging storage
% devices. Minerva \cite{minerva} extends the conventional SSD architecture
% using a FPGA-based storage controller to offload data or I/O intensive
% application code to the SSD to accelerate data analysis. However, these
% solutions cannot deal with coupled application workflows, where the data
% exchange and access patterns can be both complex and dynamic. Furthermore,
% some of these solutions require hardware modifications or special access
% priorities to the HPC systems, which can limit their use to specific
% resources.
\subsubsection{Migration}
\label{sec:migration}

\paragraph{Background:} 
Once the system has completed the initial data placement with minimal
overhead to the application, it is up to the storage system to manage the
migration of data to other storage servers and across the hierarchy. An
important reserach aspect here is how a liberated Hierarchical Storage
Manager (HSM) designed for online use, such as Sirocco, might be employed in
providing support for the migration of data. Sirocco autonomously groups the
distributed set of media and service nodes into like groups and manages the
movement of data between these groups. The grouping is managed using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

Grouping into pseudo-tiers is not a formal, nor rigid, concept within
Sirocco, at present. In classic HSM the tiering is often a formally
expressed concept, embodied both in the architecture and the implementation.
In Sirocco, though, it is not directly expressed. Instead, a client or
server that attempts to make a decision relative to migration and staging
uses these attributes to constrain the choice of candidates. For instance, a
server acting as an off-node writeback cache for a compute node might only
have access to volatile dynamic RAM for use as storage media and a client
might deposit data that is marked for persistent media as a resilience
constraint. At some point, then, such a server must copy or move the data to
another Sirocco server with media matching, or exceeding, the resilience
constraint. At that point in time, all servers managing appropriate
persistent media are candidates but second-order attributes, such as latency
and bandwidth of a candidate's managed media would tend to dominate the
ranking. Thus, the concept of a ``tier'' becomes an artifact of decision
rather than a formal concept.

Give these design decisions, our research in this project will study how
migration can serve to manage storage resident data within the constraints
defined by the application and the user.

\paragraph{Migration Policies:}
The policies that govern the migration of data across groups and across
tiers have similar challenges as initial data placement. The migration
policy must encapsulate, at a high level, the constraints defined by the
application, but must also manage resources on each server for capacity and
load reasons. We will explore policy formulations that allow \Sir to move
data within a tier to continue to meet service objectives. Another component
of migration is the temporal function that determines the utility of data
through its life time on storage. If they utility decays over time, for
example, the initial constraints of data (such as resilience requirements)
will be modified and allow more leeway in how the system handles migration.
However, as data is widely distributed the scalability of this approach must
be investigated. 
We will explore how the utility function can be used to modify migration
parameters in a scalable manner, without requiring a centralized authority
to make decisions.

\paragraph{Bounded estimates for lookup:}
One challenge for storage systems with continous decentralized migration is
the timeliness of lookup when applications issue read calls. This is due to
how no central directory exists for where data is at anytime. There are
solutions to this problem that utilize lookup tables which are updated
whenever a server initiates a migration process. However, these solutions
are not scalable. In order to have predictable performance the lookup
process must complete in a bounded time. Another solution here is to use a
distributed hashing function to determine exactly where data is placed,
reducing lookup to an $O(1)$ operation. We explore the discovery of data in
\S~\ref{sec:naming-discovery}. To support this discovery process, we will
investiagte how the migration of data can be performed in a way that enables
bounded estimates about time to first byte for a potential read operation.

% We will explore the use of a  hashing function
% that will leverage determinism to initiate the lookup process. If at most
% $k$ migration steps can happen 
% Our research intent involves short term and long term activities. In the short
% term we will modify the relevant clients to provide hit and miss related data
% and associated activities and explore an enhancement to our storage servers
% that would provide some, hopefully tightly, bounded estimate about time to
% first byte for a potential read operation.

% Further complication arises when a Sirocco server is forced to make choices for
% capacity reasons. The Sirocco architecture dictates that a server function as a
% victim cache. As media managed by a server fills the server will begin to
% manage for capacity. In order to do so, currently, it need only verify or make
% a copy of the data on another server that meets the resiliency constraint of
% the held data prior to removing the local copy and in the current
% implementation the server makes no attempt either to remember it ever had the
% data or to notify others about what it is doing.  others that it has removed
% the local copy or remember that it has

% Taken together then, the above at least complicates the higher level research
% questions discussed in this proposal, potentially even renders some
% unachievable since a few involve determinism. We are forced, then, to consider
% either or both mitigating management policies within and between Sirocco
% servers as well as explicitly allowing clients to participate in capacity
% management decisions, for a time at least, on the relevant Sirocco servers.

% \paragraph{Migration Policies:}
% In the long term we hope that an
% augmented client-server API providing usage hints will allow us to modify local
% server policy algorithms to more effectively manage it's media and communicate
% to other servers holding copies of the relevant data how they might best manage
% theirs. In the event that this approach proves ineffective at meeting the
% overall approach described in this proposal we will explore modifying Sirocco
% servers to provide notifications about locations of the relevant data as they
% migrate and stage as well as supporting client-directed pinning of location,
% for a period at least. Client-directed pinning could introduce potentially
% serious negative impacts to Sirocco's overarching goals though, since the
% fundamental design of Sirocco assumes much about opaque and unrestricted data
% movement, so this is contemplated only as a last resort.

\subsubsection{Making data avilable on multiple tiers}
\label{sec:manag-data-mult}

\paragraph{Background:} Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is
There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
full resolution data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
smaller data sets on the higher tiers for faster access.

There are a
number of drivers that push approaches to one or the other extreme:
(1) if a high tier has sufficient room for an application's working
set, approaches that move the working set into that tier will perform
well; Conversely, if a high tier very rarely captures working sets,
it is better to fill it with auxiliary data that speeds up access
to the working set captured by a lower tier; working sets only
reduce capacity misses but the concept can be generalized to the
likelihood of having the right data in place, i.e. proactively
moving data to high tiers accessed in predictable patterns reduces
compulsory misses as well; (2) if the application generates parallel
read/write requests and strong consistency semantics are important,
approaches that use auxiliary data to speed up access to persistent
shared data might yield better performance than approaches that
require coherence overhead to keep all copies on higher tiers
consistent; (3) if the application's access patterns are mismatched
with low tier access characteristics, a combination of approaches
that use auxiliary data and move data between tiers in order to
convert access patterns into a better match will perform well.

% Any other drivers?
\paragraph{Thinning of Data:} 
This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so. We will explore how the storage system can perform these
operations to meet the service level objectives for each use case, while
optimizing the overall utility from the system perpective.

@@

\subsubsection{Purging data from the system}
\label{sec:purging-data}
(only outline since I have to go)
1. Purging is important from the perspective of preventing uncontrolled
growth of data on a system
2. Purging is hard because no one wants to lose data and we need to explore
the policies that govern it
3. If we have a good utility function and a good way to refactor the data we
could purge the ``less important'' pieces and maintain a high level of
knowledge without requiring very high amount of storage
3. Purging needs to happen in a decentralized way but global utility has to
be maximized - thats a pretty hard problem

% \subsubsection{Preliminary work and results} 
% \label{sec:prework}

% Our recent work \cite{tongipdps15}
% explored a two-tiered staging method that spans both DRAM and solid state
% disks (SSD). It allows us to support both code coupling and data management
% for data intensive simulation workflows in a loosely coupled manner. In
% addition, our application-aware adaptive data placement has demonstrated the
% effectiveness of using user provided access information as ``hints'', and
% runtime data access pattern history to improve data access efficiency and
% overall end-to-end execution. Besides, additional data placement
% optimization \cite{qiansc15} that leverages physical network topology along
% with data access patterns has also demonstrated lower data access costs and
% good scalability. Our work have been deployed and proven with real
% applications, e.g., coupled combustion simulations (DNS-LES) that are part
% of the ExaCT Exascale Co-design Center and coupled Fusion simulations
% (XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems
% such Titan at ORNL.
